---
title: "Simple Time Series Analysis Example"
author: "Venkat"
date: "3/25/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

Simple Analysis of the AirPassengers dataset in R.

```{r}
library(IRdisplay)
library(magrittr)
library(scales)
library(gridExtra)
library(forecast)
library(tseries)
library(ggthemes)
library(ggplot2)
```

## Load the data 



```{r}
data(AirPassengers)
head(AirPassengers);dim(AirPassengers)

plot(AirPassengers)
```
Note that the Plot shows a definite cyclycal pattern


## Check for Stationarity
Option 1: Use ACF Plots
```{r}
g1 <- ggAcf(AirPassengers,type="correlation") + ggtitle("Autocorrelation Function ACF") + xlab("Lag") + ylab("ACF")
g2 <- ggAcf(AirPassengers,type="partial") + ggtitle("Partial Autocorrelation PACF Plot")  + xlab("Lag") + ylab("PACF")
grid.arrange(g1,g2)
```
Option 2: Use Augmented Dickie Fuller Test
```{r}
adf.test(AirPassengers)
```
Note a p-value of 0.01 indicates we REJECT the NULL Hypothesis that the data is Non-Stationary.

I.E. (double negatives!!!) - the data is STATIONARY



## Fitting models

Will try 3 models ARMA, ARIMA and STL.
Expecting that the ARIMA model will not do so well - since the data is Stationary, theoretically no Differencing should be needed. Interested to see how STL stacks up with ARMA



```{r}
#fit ARMA model
arma.model <- auto.arima(AirPassengers, max.d=0)
arma.model

#fit ARIMA model
arima.model <- auto.arima(AirPassengers)
arima.model

#fit STL model
  #1 - first transform the data to a time series
ap.ts <- ts(AirPassengers,frequency=12)

  #2 - fit a model
stl.model <- stl(ap.ts,s.window="periodic")
autoplot(stl.model)

```

## Plot residuals for ARMA and ARIMA model
```{r}

arma.resid <- resid(arma.model)
arima.resid <- resid(arima.model)


g1 <- autoplot(arma.resid)
g2 <- autoplot(arima.resid)
grid.arrange(g1,g2)
```

## Compare the forecasts of each model

```{r}
#forecasting for 2 years. 80% confidence
arma.forecast <- forecast(arma.model,h=24,level=80)
arima.forecast <- forecast(arima.model,h=24,level=80)
stl.forecast <- forecast(stl.model, h=24, level=80)

g1 <- autoplot(arma.forecast)
g2 <- autoplot(arima.forecast)
g3 <- autoplot(stl.forecast)
grid.arrange(g1,g2,g3)
```
## Conclusion
No difference between the ARMA and ARIMA Models (expected as the data is stationary)
STL performed poorer that I had expected
